---
title: "R Notebook"
output: html_notebook
---
## Import libraries
```{r}
library(dplyr)
library(stringdist)
```


## Importing the notebook
```{r cache = TRUE}

stormData <- read.csv("StormData.csv.bz2")

```
```{r cache = TRUE}
subData <-  stormData[c("EVTYPE" , "FATALITIES", 
                        "INJURIES","PROPDMG", 
                        "PROPDMGEXP",
                        "CROPDMG","CROPDMGEXP")]
subData$EVTYPE = toupper(subData$EVTYPE)
subData$PROPDMGEXP = toupper(subData$PROPDMGEXP)
subData$CROPDMGEXP = toupper(subData$CROPDMGEXP)
```
For the *EXP values, they stand for the exponents of the values.  
"K" is 10^3  
"M" is 10^6
"B" is 10^9
```{r}
exponent.list = list("K" = 1000, "M" = 10^6,
                     "B" = 10^9, "0" = 1)
names(exponent.list)

```


```{r}
unique(subData$PROPDMGEXP)

```
The PROPDMGEXP not in exponent.list have some PROPDMG values not equal to 0. However, it is impossible to determine what the invalid PROPDMGEXP values are supposed to be. I replaced the PROPDMGEXP value where PROPDMG is 0 with "0". I  also replaced "" in PROPDMGEXP with "0"
Since the rest rows with PROPDMGEXP not in exponent.list are about 42 which is about 0.4%, they can be deleted  
```{r}
subData <- subData %>% 
    mutate(PROPDMGEXP = ifelse(PROPDMG == "0" | PROPDMGEXP == "",
                               "0", PROPDMGEXP) ) %>%
    filter(PROPDMGEXP %in% names(exponent.list) )
```

```{r}
unique(subData$CROPDMGEXP)
```
The rows with CROPDMGEXP not in exponent.list have a CROPDMG value of 0 and can be replaced with "0".
```{r}
subData <- subData %>% 
    mutate(CROPDMGEXP = ifelse(CROPDMGEXP %in% names(exponent.list),
                               CROPDMGEXP,"0"))
```

## To calculate the Damage number with DMG and DMGEXP and find total of damages
```{r}
subData <- subData %>% 
    mutate(PROPDMGNUM = PROPDMG * exponent.list[PROPDMGEXP][[1]],
           CROPDMGNUM = CROPDMG * exponent.list[CROPDMGEXP][[1]],
           TOTALDMG = PROPDMGNUM + CROPDMGNUM)
```

EVTYPE has 48 standard values as seen in page 6 
```{r}
EVTYPES<-"ASTRONOMICAL LOW TIDE
AVALANCHE
BLIZZARD
COASTAL FLOOD
COLD/WIND CHILL
DEBRIS FLOW
DENSE FOG
DENSE SMOKE
DROUGHT
DUST DEVIL
DUST STORM
EXCESSIVE HEAT
EXTREME COLD/WIND CHILL
FLASH FLOOD
FLOOD
FROST/FREEZE
FUNNEL CLOUD
FREEZING FOG
HAIL
HEAT
HEAVY RAIN
HEAVY SNOW
HIGH SURF
HIGH WIND
HURRICANE (TYPHOON)
ICE STORM
LAKE-EFFECT SNOW
LAKESHORE FLOOD
LIGHTNING
MARINE HAIL
MARINE HIGH WIND
MARINE STRONG WIND
MARINE THUNDERSTORM WIND
RIP CURRENT
SEICHE
SLEET
STORM SURGE/TIDE
STRONG WIND
THUNDERSTORM WIND
TORNADO
TROPICAL DEPRESSION
TROPICAL STORM
TSUNAMI
VOLCANIC ASH
WATERSPOUT
WILDFIRE
WINTER STORM
WINTER WEATHER"

EVTYPES<-unlist(strsplit(EVTYPES,split = "
"))

#add some more
EVTYPES <- c(EVTYPES, c("WIND",""))
```

To clean EVTYPES
```{r}
uniqueTYPES <-  unique(subData$EVTYPE)

subData$EVTYPE<-gsub("TSTM","THUNDERSTORM",subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("FLD","FLOOD",subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("^FREEZE$","FROST/FREEZE",subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("^FROST$","FROST/FREEZE",subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("^STORM SURGE$","STORM SURGE/TIDE",subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("^TIDE$","STORM SURGE/TIDE",subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("^TYPHOON$","HURRICANE (TYPHOON)",subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("^HURRICANE$","HURRICANE (TYPHOON)",subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("^EXTREME COLD$","EXTREME COLD/WIND CHILL",
                     subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("^EXTREME WIND CHILL$","EXTREME COLD/WIND CHILL",
                     subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("^COLD$","COLD/WIND CHILL",
                     subData$EVTYPE,fixed = T)
subData$EVTYPE<-gsub("WIND CHILL$","COLD/WIND CHILL"
                     ,subData$EVTYPE,fixed = T)

uniqueEVTYPE <-  unique(subData$EVTYPE)

#Perform fuzzy text search
d <- expand.grid(uniqueEVTYPE,EVTYPES)  # Distance matrix in long form
names(d) <- c("EVTYPE.DF.COLUMN","EVTYPE.CATEGORIES")
d$dist <- stringdist(d$EVTYPE.DF.COLUMN,d$EVTYPE.CATEGORIES,
                     method="jw") # String edit distance 

fuzzyResult <- do.call(rbind, unname(by(d, d$EVTYPE.DF.COLUMN, function(x) 
    x[x$dist ==min(x$dist),])))     
```

